---
title: "VisualizeSeurat"
author: "JuHyunJeon"
date: "2023-05-24"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Viusualize with Seurat

In this document, I will go through the basic workflow of Seurat but try to explore some alternative visualization techniques that can be employed to make graphs more readable.

## Import libraries and download data

```{r cars}
library(tidyverse)
library(dplyr)
library(Seurat)
library(SeuratData)
library(ggplot2)
library(patchwork)

# I will use pbmc3k for his dtaset.
InstallData("pbmc3k")

data("pbmc3k.final")
pbmc3k.final$groups <- sample(c("group1", "group2"), size = ncol(pbmc3k.final), replace = TRUE)
features <- c("LYZ", "CCL5", "IL32", "PTPRCAP", "FCGR3A", "PF4")
pbmc3k.final
```

## Ridgeline plots

The following is a sample plot from the vignette.

```{r seurat-base-ridgeplot, fig.width = 10, fig.height = 10, dpi = 450}
RidgePlot(pbmc3k.final, features = features, ncol = 2)
```

However, this is not the best visualization technique for this data for several reasons.

- There is no order to the identities.
- The plot doesn't convey more information being a ridgeline plot. However, there is no clear benefit of overlapping.
  - For LYZ, we do not know the expression level shape of NK around expression level 3.
  - For PF4, the graph gives no information other than Platelet having outliers which doesn't necessarily have to be in ridgeline plot format.
- Still for LYZ and CCL5, it shows a collection of distribution of gene expression levels.

## Violin Plot

```{r, seurat-base-vlnplot, fig.width = 10, fig.height = 10, dpi = 300}
VlnPlot(pbmc3k.final, features = features)
```

- Violin plot suits the dataset better than ridgeline plot since it doesn't hide any distribution.
```{r}
df <- data.frame(seurat_annotations = character(), features = character(), value = numeric())
uniq_annot <- unique(pbmc3k.final$seurat_annotations)
for (annot in uniq_annot){
  pbmc.subset <- subset(pbmc3k.final, subset = seurat_annotations == annot)
  seurat_df <- GetAssayData(pbmc.subset)[features, ]
  for (f in features){
    for(i in 1:length(seurat_df[f, ])){
      df <- df %>% add_row(seurat_annotations = annot, features = f, value = seurat_df[f, i])
    }
  }
}

df$features <- factor(df$features,levels = c("LYZ", "CCL5", "IL32", "PTPRCAP", "FCGR3A", "PF4"))
```


```{r, fig.width = 18, fig.height = 12}
ggplot(df, aes(x = seurat_annotations, y = value)) + geom_boxplot() + facet_wrap(~features, nrow = 2, ncol = 3) + theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))
```

